Structured Synchronous Reactive Programming for Game Development \\ \Large{Case Study: On Rewriting Pingus from C++ to \CEU}

### 2017-05-23 paper.tex
...

In this work, we advocate structured synchronous reactive programming as an
expressive and productive alternative for game logic development.
We present a case study of rewriting Pingus from C++ to \CEU with the
contributions as follows:
%
\begin{itemize}
\item Identifying five recurrent control-flow patterns that likely apply to
      other games:
        \emph{Finite State Machines},
        \emph{Continuation Passing},
        \emph{Dispatching Hierarchies},
        \emph{Lifespan Hierarchies},
        \emph{Signaling Mechanisms}.
\item Applying idiomatic code in \CEU as alternative solutions for six selected
      behaviors in the game logic.
\item Presenting an in-depth qualitative analysis of the proposed solutions in
      comparison to the original implementations in C++.
\end{itemize}
%
A control-flow pattern is a recurring technique to describe dependencies and
explicit orders between statements (or groups of statements) in a program.
For instance, consider how a key press stimulus propagates through the game
entities and also what happens with them if the stimulus causes the end of the
game.
%
\CEU supports primitives that help describing complex control-flow
relationships in the game logic more concisely.
%
The rewriting process consisted of identifying sets of callbacks implementing
control flow in the game and translating them to \CEU using appropriate
structured constructs.
%
As an example, a double mouse click is characterized by a first click, followed
by a maximum amount of time, followed by a second click.
This behavior depends on different events (clicks and timers) which have to
occur in a particular order.
In C++, the implementation involves callbacks crossing reactions to successive
events which manipulate state variables explicitly.
%
Although we documented in detail two use cases for each of the five patterns,
in this work we only present six cases due to space constraints.
%
This work focuses on a qualitative analysis for the programming techniques
that we applied during the rewriting process.
Not all techniques result in reduction in LoCs (especially considering the
verbose syntax of \CEU), but have other effects such as eliminating shared
variables and dependencies between classes.

The rest of the paper is organized as follows:
Section~\ref{sec.codebase} gives an overview of the Pingus codebases in C++ and
\CEU.
Section~\ref{sec.patterns} describes the five control-flow patterns and
discusses six case studies in detail.
Section~\ref{sec.related} discusses related work.
Section~\ref{sec.conclusion} concludes the paper.

@@

@@

@@

use cases first -> patterns later

We can identify control-flow behaviors in C++ by looking for class members with
identifiers resembling verbs, statuses, and counters (e.g.,
\code{pressed},
\code{particle\_thrown},
\code{mode}, and
\code{delay\_count}).
Good chances are that variables with these ``suspicious names'' encode some
form of control-flow progression that cross multiple callback invocations.

We selected 6 representative game behaviors and describe their
implementations in C++ and \CEU.
We also categorize these behaviors in 5 abstract control-flow patterns that
likely apply to other games:

@@

Some previous work~\cite{a,b,c} discuss how the GoF patterns~\cite{gof} apply
to video game development.
    - generally?
        - mostram exemplos abstratos
            - imagine um jogo...
                - assim poderiamos aplicar...
            - sem codigo real
            - nao aplicam a jogos existentes

>>>>

In this work, we advocate structured synchronous reactive programming as an
expressive and productive alternative for game logic development.
We present a case study of rewriting Pingus from C++ to \CEU with the
contributions as follows:
%
\begin{itemize}
\item Applying idiomatic code in \CEU as alternative solutions for six selected
      behaviors in the game logic.
\item Presenting an in-depth qualitative analysis of the proposed solutions in
      comparison to the original implementations in C++.
\item Identifying five recurrent control-flow patterns that likely apply to
      other games:
        \emph{Finite State Machines},
        \emph{Continuation Passing},
        \emph{Dispatching Hierarchies},
        \emph{Lifespan Hierarchies},
        \emph{Signaling Mechanisms}.
\end{itemize}
%
A control-flow pattern is a recurring technique to describe dependencies and
explicit orders between statements (or groups of statements) in a program.
For instance, consider how a key press stimulus propagates through the game
entities and also what happens with them if the stimulus causes the end of the
game.
%
\CEU supports primitives that help describing complex control-flow
relationships in the game logic more concisely.
%
The rewriting process consisted of identifying sets of callbacks implementing
control flow in the game and translating them to \CEU using appropriate
structured constructs.
%
As an example, a double mouse click is characterized by a first click, followed
by a maximum amount of time, followed by a second click.
This behavior depends on different events (clicks and timers) which have to
occur in a particular order.
In C++, the implementation involves callbacks crossing reactions to successive
events which manipulate state variables explicitly.
%
Although we documented in detail two use cases for each of the five patterns,
in this work we only present six cases due to space constraints.
%
This work focuses on a qualitative analysis for the programming techniques
that we applied during the rewriting process.
Not all techniques result in reduction in LoCs (especially considering the
verbose syntax of \CEU), but have other effects such as eliminating shared
variables and dependencies between classes.

The rest of the paper is organized as follows:
Section~\ref{sec.codebase} gives an overview of the Pingus codebases in C++ and
\CEU and describes our approach to identify and rewrite the control flow in the
game.
Section~\ref{sec.patterns} discusses six case studies in detail which are
categorized in five control-flow patterns.
Section~\ref{sec.related} discusses related work.
Section~\ref{sec.conclusion} concludes the paper.

@@

As a general rewriting rule, we could identify control-flow behaviors in C++ by
looking for class members with identifiers resembling verbs, statuses, and
counters (e.g.,
\code{pressed},
\code{particle\_thrown},
\code{mode}, and
\code{delay\_count}).
Good chances are that variables with these ``suspicious names'' encode some
form of control-flow progression that cross multiple callback invocations.

During the course of the rewriting process, we could identify more general
control-flow patterns which likely apply to other games as well.

@@

In this section, we select six representative game behaviors and describe in
detail their implementations in C++ and \CEU.
%
We also categorize these behaviors in five abstract control-flow patterns as
follows:

@@

@@

The control-flow patterns closely relate to the \emph{GoF} behavioral
patterns~\cite{gof}, which some previous work discuss in the context of video
game development~\cite{a,b,c}.
%
different
    - generally?
        - mostram exemplos abstratos
            - imagine um jogo...
                - assim poderiamos aplicar...
            - sem codigo real
            - nao aplicam a jogos existentes
    - propose solutions
        - eliminate
for instance, the \emph{state pattern}, state machines, sequences
trivial case of the bomber animation is a series of statements separated
by await statements
visitor, adopted in the language
